To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis

### 2020-04-07 Thesis.tex
Add files via upload
Thesis

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis}
\date{}
\author{Danil Fedchenko\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
\begin{document} % 
	\maketitle
	\begin{abstract}
		Many e-commerce platforms that connect buyers and sellers employ recommender systems to help customers find products and services. These systems may have several potential effects on strategies of competing sellers: on one hand, more people become aware of products, so sellers may sell more goods and benefit from the presence of the recommender system. On the other hand, at the same time, more people become aware of the competitor seller. The latter effect may intensify the price competition and may hurt sellers but, of course, benefit customers. However, many platforms charge firms for sales and as a result have the stakes in firms profits. Thus, by designing the recommender policy the platform should balance the pro-competitive care about customers and anti-competitive incentives to keep high prices. The main goal of this work is to study this trade-off in details. I am solving for an optimal recommender policy and investigate factors that may affect it.
	\end{abstract}
\newpage
\tableofcontents
\newpage
	\section{Introduction}
	\section{Literature review}

@@

	One of the first papers which highlights the issue from the point of view of Economics is work by \cite{avery1999market}. \cite{avery1999market} considered the influence of the recommender system on choice regarding the consumption of good for the sequentially arriving agents. They model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review. The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.

	\cite{bergemann2006optimal} address the question of optimal pricing under the presence of recommender system. They consider the seller with recommender system who is competing for unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system. The seller with recommender system uses reviews from past customers to make recommendations to the newcomers. The authors consider several type of equilibria concluding that the presence of the recommender system creates an additional value by reducing uncertainty for agents about the quality of the good.

@@

	Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
	\cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes. They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between. There is also a recommender system that receives a noisy information about each customer's location and can recommend one of the sellers to the customers. Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers. In particular, they are looking at prices and advertising expenditure. The main finding of this paper is that the presence of the recommender system may fundamentally change the nature of competition. In particular, the authors demonstrate that with the recommender system prices and advertising expenditure switch from being strategic substitutes to strategic complements. This results in the fact that with recommender system sellers prefer to advertise less, relying more on advertising provided by the recommendation system (the authors called this effect \textit{advertising effect}). At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition. Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers becomes worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.

	Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper). In the first one, considering the same set-up as \cite{chen2016advertising}, the authors do not allow sellers to advertise themselves, assuming the shares of informed and uninformed customer to be exogenous. The authors demonstrate that the presence of the recommender system on one hand increase the share of informed customers and hence increases demand on goods, but on the other hand it intensifies competition by making a demand more elastic. \cite{li2018recommender} refer to these two effects as \textit{demand effect} and \textit{substitution effect} respectively. In the similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of recommender system on sellers becomes a result of interaction between these tow effects. Also, in this paper the authors make the first attempt to endogenise the recommender policy, showing that the recommender system may increase its profit by changing the recommendation strategy. In the second version of the paper the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement. In the two (in fact, three) papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.

	\cite{levin2009quality} add a vertical component, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system. In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.

	\section{Model}
	The model I consider is similar to that considered by \cite{levin2009quality}, the difference is that I introduce a new player -- a platform which employs a recommender system (I will use words ``platform'' and ``recommender system'' interchangeably).

	There is a unit mass of heterogenous customers. Preferences of customer are characterized by the location of the customer on the unity interval. Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0. Goods of each seller may have different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers. Sellers are competing in prices which they set simultaneously. In addition, the sellers may costly disclose the information about the quality of their good prior to price setting. The recommender system observes prices and sellers' decisions regarding the quality disclosure, and makes recommendation to customers which seller to choose. In the baseline model I assume that the recommender system knows the exact location of the customer, lately this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations. Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.

	The solution concept is perfect Bayesian equilibrium, and all parties are assumed to be risk-neutral.

	The next three subsections describes formally all parties of my model.

@@

	Prior to price-setting sellers privately learns the quality of their good and should decide whether to disclose this information to the platform or not. Upon disclosure the sellers bear cost $c$. Moreover, the sellers pay the platform commission from sales. That is, the seller's $i$ profit is given by the following formula
	$$\pi_i = (1-\alpha)p_i D_i - c d_i$$
	where $$d_i = \begin{cases}
	1, \text{ if sellers }i\text{ discloses }\\
	0, \text{ otherwise }
	\end{cases}$$ and $D_i$ is the demand on good $i$ which may be a function of prices and decisions regarding the information disclosure, and is determined by the recommender policy the platform employs.

@@

	\subsection{Timing}
	The timing of the game looks as follows:
	\begin{enumerate}
		\item The platform commits to the commission rate $\alpha$ and the recommender policy.
		\item The sellers privately observes $q_A, q_B$ and decide whether to disclose it to the platform.
		\item The sellers simultaneously set prices.
		\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that the payoffs are realized and the game ends.
	\end{enumerate}

	\section{Results}
	Firstly, I consider the case with observable qualities and without the recommender system. I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects equilibrium behavior of sellers.

@@

	 Given the demand, 	sellers set prices $p_A, p_B$ simultaneously to solve the following problems:
	 $$(1-\alpha)\cdot p_A \cdot \hat{x} = p_A \left(\frac{1}{2} - \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}\right) \to \underset{p_A}{\max}$$
	 $$(1-\alpha)\cdot p_B \cdot (1-\hat{x}) = p_B \left(\frac{1}{2} + \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} - \frac{q_A - q_B}{2}\right) \to \underset{p_B}{\max}$$
	 The system of first order conditions is given by:
	 \begin{align*}
	 \begin{cases}
	 p_A = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_B}{2} + \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}\\
	 p_B = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_A}{2} - \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}
	 \end{cases}
	 \end{align*}
	 Solving the system above yields $p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_i - q_{-i})}{3(\gamma - \alpha)}$
	\end{proof}
	Comparing the results of the Proposition \ref{with} with Lemma \ref{without} several things are worth noting. First of all, with the recommender system the term $\frac{p_A-p_B}{2}$ in the expression for demand has an extra multiplier $\frac{\gamma-\alpha}{\gamma} \in (0, 1)$. The latter fact means that is the seller $A$ decides to undercut the competitor and sets a price $p_B - \varepsilon$, with the recommender system the resulting gain in demand will be $\frac{\gamma - \alpha}{\gamma} \frac{\varepsilon}{2}$ which is strictly below $\frac{\varepsilon}{2}$ the result we obtained for the case without the platform. That means that now the sellers have fewer incentives to undercut the competitor, i.e. the presence of recommender system which has stakes in sellers' profits softens competition. As a result, the equilibrium prices a higher. The picture below (Fig. \ref{fig1}) demonstrates graphically the impact of the recommender system.

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{ImpactRS}
		\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
	\end{figure}
	As we can see, there exists a share of customers $\Delta = \frac{\alpha}{2 \gamma} (p_B - p_A)$ such that they strictly prefer the firm $A$ to the firm $B$ while are recommended the firm $B$. This happens since from the platform's standpoint the losses in consumer surplus are lower than gains driven by high-price sales. Note, also, that $\frac{\partial \Delta}{\partial \alpha} > 0$, $\frac{\partial \Delta}{\partial \gamma} < 0$ and $\lim_{\gamma \to \infty} \Delta = 0$. That is, the impact of the platform eliminates completely if $\gamma$ becomes very high, and the more the platform is profit-oriented (the higher is $\alpha$) the higher is $\Delta$.
	\subsection{Qualities are not observable}
	Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decides to disclose it. In this subsection I derive the optimal recommender system of the platform. Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do no not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.

	The optimal platform's policy can be found separately: firstly the platform may obtain the optimal recommendation (as a function of $\alpha$) treating the commission rate $\alpha$ as given, and then find an optimal commission rate. The next proposition characterize the optimal recommendation strategy given the commission rate $\alpha$.

@@

	From the Bayes rule it follows that if the seller does not disclose the information the platform forms beliefs that the quality is $\frac{q^*}{2}$. Given the platform's beliefs regarding the quality, the proof exactly repeats that of Proposition \ref{with}. It remains to find $q^*$. This threshold is found as the minimal quality that renders the seller indifferent between disclosing and not disclosing the information. That is, $q^*$ is determined as a solution to the following equation:

@@

However, it turns out that depending on cost of information disclosure, the amount of information disclosed in equilibrium depends differently on the recommender system precision.

>>>>

\documentclass[a4paper]{article}
\usepackage[12pt]{extsizes} % 
\usepackage{setspace}
\doublespacing
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
%\usepackage{harvard}
\usepackage[round]{natbib}
\usepackage{pdfpages}
\usepackage{tikz}
\usepackage{makecell}
\usepackage{amsthm}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage{breakcites}
\interfootnotelinepenalty=10000
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
%\newenvironment{abstract}[0]{\small\rm
%	\begin{center}ABSTRACT
%		\\ \vspace{8pt}
%		\begin{minipage}{5.2in}\smalllineskip
%			\hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathcal{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=25mm, top=20mm, right=25mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{definition}[theorem]{Definition}
%\let\cite = \citeasnoun
\title{To run with the hare and hunt with the hounds: Optimal recommendation system with competing sellers. \\ Master Thesis}
\date{}
\author{Danil Fedchenko}%\thanks{I am very grateful to Sergei Izmalkov for his help during the work on this project}}
\begin{document} % 
	\maketitle
	\begin{abstract}
		Many e-commerce platforms that connect buyers and sellers employ recommender systems to help customers find products and services. These systems may have several potential effects on strategies of competing sellers: on one hand, more people become aware of products, so sellers may sell more goods and benefit from the presence of the recommender system. On the other hand, at the same time, more people become aware of the competitor seller. The latter effect may intensify the price competition and may hurt sellers but, of course, benefit customers. However, many platforms charge firms for sales and as a result have the stakes in firms profits. Thus, by designing the recommender policy the platform should balance the pro-competitive care about customers and anti-competitive incentives to keep high prices. The main goal of this work is to study this trade-off in details. I am solving for an optimal recommender policy and investigate factors that may affect it.
	\end{abstract}
\newpage
\tableofcontents
\newpage
	\section{Introduction}

		Suppose you came to New-York and would like to find a good restaurant to go out tonight. Or you are going to spend your upcoming vacation on the coast of Spain and you need to find a hotel. What would you do? What will be the first thing you interact with?  Before the World Wide Web, people in such cases relied upon guide-books and brochures distributed by travel agents. Those guides usually quickly became out-of-date and were criticized for containing misleading information.

	After the advent of the Internet, finding up-to-date information has become much easier. To make this process even easier Internet platforms began to appear. These platforms aggregated customers' reviews on products and service providers. Some of those platforms simply collect the reviews and present a rating that is just an average measure of consumers' satisfaction (weighted or simple mean of rates). Examples include \textit{Yelp} for restaurants, home and car services, or \textit{SiteJabber} for online businesses.

Other platforms came further and began to not only collect the reviews but also to make individual recommendations based on these reviews. Usually, these recommendations take a form of ranking of several alternatives that are ranked according to the secret propriety algorithm that the particular platform employs. The most widely known examples of such a kind of platforms include: \textit{TripAdvisor} which presents a ranking on Hotels and Restaurants based on previous users' experience; \textit{Trustpilot} which is one of the most popular review services for a large variety of businesses; \textit{Glassdoor} that uses reviews of employees to rank companies by workers' satisfaction; and many other examples.

In the rows of examples, the special attention should be paid to online retailers such as \textit{Amazon}, \textit{eBay}, \textit{AliExpress} and many others. These platforms stand at the forefront of online Economics, generating huge amounts of money and completely transforming the shopping experience for people all other the world. These platforms play roles of intermediaries between buyers and sellers, helping the buyers to find the goods they need. And this process clearly includes making recommendations which seller to display you when you are searching for something.

	All these services discussed above use some variety of what is called \textit{recommender system} which may be defined as any system which uses information about past users' experience in order to make the recommendations to the newcomers. These recommendations may be direct or indirect (e.g. the platform can provide a ranking of alternatives, this is not the direct recommendation to choose the top-ranked alternative, however, users understand that the platform thinks that this alternative is the best fit for them, so the platform implicitly recommends this alternative).

I argue, that nowadays the recommender systems are one of \textbf{the key elements} of online Economics. Any Internet or mobile service you can think of almost for sure uses some form of a recommender system. That is why it is crucially important to understand how to recommend optimally and what factors should be taken into account once the service is designing its recommender system.

For example, consider a retailer (such as \textit{Amazon} or \textit{eBay}), these retailers charge sellers commissions for each sale. These commissions usually take the form of a percentage commission. That means that the retailer obtains a higher profit if the price of the deal is higher. So, one of the questions that could potentially arise then is: should, in this case, the recommender system which Amazon employs out of two sellers who sell (almost) substitutable goods recommend the one who has a higher price? Or suppose that \textit{eBay} is thinking on the commission rate it charges (now it is about 8\%), should the commission be set at the highest possible level that guarantees that sellers do not refrain from using the platform? What if, at the same time, the quality of goods may differ and the sellers may have an option to spend some resources on disclosing this quality (e.g. to make the detailed description or take high-resolution photos). Clearly the higher is the commission the less willing are the sellers to spend money on quality disclosure: cost does not outweigh benefits. But if sellers do not disclose the information this may adversely affect the customers' satisfaction since this may create a mismatch between buyers and sellers: people end up with the good which does not fit them well. Thus, thinking about commission rate the platform should take this aspect into account. Additionally, suppose that the recommendation system is not precise i.e. it sometimes makes mistakes on figuring out the customer's preferences. How does this inaccuracy affect the behavior (price setting) of competing sellers and how does it interact with their incentives to disclose the quality of their goods? What should be the optimal commissions and recommender policy in this case? Although in this work I am touching upon all questions mentioned above, the main objective of this paper is to understand what should be the optimal recommendation strategy and commission rate the platform charges the competitive sellers, with the particular emphasis on the interaction with sellers' incentives to disclose private information about their goods.

	To address these questions I build a theoretical one-period model (please see the detailed description of the model in the corresponding section). I consider the unit mass of heterogeneous customers whose preferences are summarized by the point on this unit line. At the extremes of the unit interval, there are two sellers. Customers demand one good and derive utility upon consuming the good but bear ``transportation'' cost by purchasing. This cost, of course, has little to do with any transportation but rather represents the outside option, opportunity cost of the customer: the utility the customer should refrain from in order to consume the good. I suppose that the qualities of goods offered by sellers can be different and that these qualities are \textit{ex-ante} privately known only to the corresponding sellers who can costly disclose this information before setting prices. The main player in this system is the recommender system (the platform) who beforehand receives (possibly noisy) signal regarding each customer's location and then may recommend any of the two sellers to this customer. This feature of my model has clear real-life interpretation: location of the customer represents his or her tastes, the signal which reveals the location can be interpreted as the result of some fancy machine-learning algorithm the majority of platform nowadays use in order to assess, to estimate the true preferences, tastes of the customer. The platform also sets a commission rate (percentage) it charges the sellers for each sale. However, I explicitly assume that the platform cares not only about the profit (commissions) but also about customers i.e. consumer surplus directly enters the platform's objective function. I suppose that this assumption is not very far away from reality: clearly, if the platform cares only about the profit and this adversely affects customers' satisfaction and happiness from using the service, in the long-run people simply will refuse using it or switch to the competitive platforms. The latter suggests that for the long-run sustainability the platform should pay some attention to the consumer surplus.

In the work, I find the optimal recommender policy and investigate the effect of the presence of the recommender system on equilibrium behavior of sellers. I demonstrate that the recommender system which has stakes in the sellers' profits softens competition. The latter effect leads to higher prices. The amount of information disclosed in equilibrium is determined in the interplay between two effects: on one hand, higher commission directly decreases profits and, hence, incentives of sellers to disclose the information; on the other hand, the higher commission increases platform's stakes in profits and softens competition which results in higher prices and higher profits. The total value of these two effects is determined by the degree of consumer-orientedness of the platform: if it is high enough, increase in prices does not outweigh direct losses, so the total effect is negative; in the opposite case total effect is positive and an increase in commission increases the amount of information disclosed.

In the extension of the baseline model, I allow for the not perfectly precise recommender system. Namely, I assume that the signal, the platform receives reveals the true location of the customer only with some probability $\beta$, and with probability $1 - \beta$ the signal follows a uniform distribution. This has several consequences: first of all, now the platform adjusts its recommender strategy incorporating this inaccuracy. This adjustment manifests itself in the way how the platform treats and processes the signal regarding the location, now the platform takes into consideration that the signal could potentially be wrong. The latter effectively increases the share of customers in the middle, increasing incentives for sellers to undercut the competitor and to get access to this increased demand. Thus, the direct effect of the platform's inaccuracy intensifies the competition between sellers. This decreases the sellers' profits and as a result, decreases sellers' incentives to reveal the quality. At the same time, the recommender system's inaccuracy increases the attention the platform pays to the quality of the good when it makes recommendations. The latter effect increases the sellers' returns on quality disclosure making the latter more profitable. Eventually, the total effect also depends on parameters: if the cost of disclosure is high enough, the first effect dominates and improving the recommender system precision decreases the amount of information the sellers disclose, for the low cost of disclosure the second effect prevails and increasing the platform's precision increases the amount of information disclosed.

However, it turns out that increasing the precision of the recommender system up to 1 is not the optimal thing to do. The last thing I show in this work is that there exists some optimal level of $\beta \in (0, 1)$. This can be explained as follows: making a recommender system a bit imprecise, intensifies competition and lowers prices, at the same time if precision is not too low, the direct negative effect on consumer surplus is not very high. As a result, I show that some noise in the platform's evaluation of the customers' preferences can be beneficial for the platform.

The rest of the paper is organized as follows: the next section presents the literature review, Section 3 describes the model, Section 4 contains the main results, and last three Sections encompass a brief discussion, possible extensions and directions for the future research, and a short conclusion.

	\section{Literature review}

@@

One of the first papers which highlight the issue from the point of view of Economics is work by \cite{avery1999market}. \cite{avery1999market} considered the influence of the recommender system on a choice regarding the consumption of good for the sequentially arriving agents. The model they consider include sequentially arriving agents who can either consume the good and leave a review for the next agents or he can wait until somebody else will try the good and report the review. The main focus of the work is on finding a socially optimal mechanism, in particular, the authors propose a mechanism which includes monetary transfers to some agents to induce them to try the good.

\cite{bergemann2006optimal} address the question of optimal pricing under the presence of a recommender system. They consider the seller with a recommender system who is competing for a unit mass of heterogeneous buyers with the competitive fringe of sellers who do not deploy the recommender system. The seller with the recommender system uses reviews from past customers to make recommendations to the newcomers. The authors consider several types of equilibria concluding that the presence of the recommender system creates additional value by reducing uncertainty for agents about the quality of the good.

@@

Three papers mentioned above although address the question on optimal recommender policy, do not take into account the strategic nature of recommended alternatives.
\cite{chen2016advertising} consider the informative role of recommender system on strategies of competing sellers and the equilibrium outcomes. They consider two sellers located at the extremes of Hotelling-like (\cite{harold1929stability}) linear city who are competing for the unit mass of customers, located in between. There is also a recommender system that receives noisy information about each customer's location and can recommend one of the sellers to the customers. Although in this paper, the recommendation strategy is exogenous, the authors consider how the presence of the recommender system affects competition between sellers. In particular, they are looking at prices and advertising expenditure. The main finding of this paper is that the presence of the recommender system may fundamentally change the nature of competition. In particular, the authors demonstrate that with the recommender system prices and advertising expenditure switch from being strategic substitutes to strategic complements. This results in the fact that with recommender system sellers prefer to advertise less, relying more on advertising provided by the recommendation system (the authors called this effect \textit{advertising effect}). At the same time, prices with recommender system are lower (the authors called this effect \textit{competition effect}) which is the result of tougher competition. Interestingly, the total effect of recommender system on sellers profits becomes a result of an interplay between \textit{competition effect} and \textit{advertising effect}: for some values of parameters the sellers become worse-off with the presence of the recommender system, \textit{competition effect} dominates, while for other values the sellers benefit from ``free'' advertising provided by the platform.

Another paper, which is closely related to the one mentioned previously is the paper by \cite{li2018recommender} (actually there are two quite different versions of this paper). In the first one, considering the same set-up as \cite{chen2016advertising}, the authors do not allow sellers to advertise themselves, assuming the shares of informed and uninformed customer to be exogenous. The authors demonstrate that the presence of the recommender system, on one hand, increase the share of informed customers and hence increases demand on goods, but on the other hand, it intensifies competition by making demand more elastic. \cite{li2018recommender} refer to these two effects as \textit{demand effect} and \textit{substitution effect} respectively. In a similar way as in \cite{chen2016advertising}, the authors demonstrate that the total effect of the recommender system on sellers becomes a result of interaction between these two effects. Also, in this paper the authors make the first attempt to endogenise the recommender policy, showing that the recommender system may increase its profit by changing the recommendation strategy. In the second version of the paper, the authors elaborate on this, formulating the recommender system's objective and finding an optimal recommender policy the recommender system should implement. In the two (in fact, three) papers mentioned above, the qualities of goods were supposed to be equal and known \textit{ex-ante} to all parties which makes the type of differentiation considered to be purely horizontal.

\cite{levin2009quality} add a vertical component, focusing on the sellers' decisions regarding the quality disclosure, although, in the framework without the recommender system. In the model which I consider, I firstly endogenise the recommendation strategy that the platform implements, and secondly, I am looking at the interplay between the recommendation policy and sellers' incentives to disclose the information about the goods' quality.	
	\section{Model}
The model I consider is similar to that considered by \cite{levin2009quality}, the difference is that I introduce a new player -- a platform which employs a recommender system (I will use words ``platform'' and ``recommender system'' interchangeably).

There is a unit mass of heterogeneous customers. Preferences of customer are characterized by the location of the customer on the unity interval. Two competing sellers (firms) which I will refer to $A$ and $B$, who locate at the extremes of the unit interval, are selling substitutable goods which they produce at equal marginal cost normalized to 0. Goods of each seller may have a different quality which means that the customers with the same ``horizontal preference'' for the two sellers may derive different utility upon consuming goods of different sellers. Sellers are competing in prices which they set simultaneously. In addition, the sellers may costly disclose the information about the quality of their good prior to price setting. The recommender system observes prices and sellers' decisions regarding the quality disclosure and makes a recommendation to customers which seller to choose. In the baseline model, I assume that the recommender system knows the exact location of the customer, lately, this assumption is relaxed, and I consider the scenario when the recommender system receives a noisy signal regarding the customers' locations. Finally, the platform charges sellers a commission for sales according to a commission rate $\alpha$ which the platform can choose.

The solution concept is the perfect Bayesian equilibrium, and all parties are assumed to be risk-neutral.

The next three subsections describe formally all parties of my model.

@@

Prior to price-setting sellers privately learns the quality of their good and should decide whether to disclose this information to the platform or not. Upon disclosure, the sellers bear cost $c$. Moreover, the sellers pay the platform commission from sales. That is, the seller's $i$ profit is given by the following formula
$$\pi_i = (1-\alpha)p_i D_i - c d_i$$
where $$d_i = \begin{cases}
1, \text{ if sellers }i\text{ discloses }\\
0, \text{ otherwise }
\end{cases}$$ and $D_i$ is the demand on good $i$ which may be a function of prices and decisions regarding the information disclosure, and is determined by the recommender policy the platform employs.

@@

	\subsection{Timing}
The timing of the game looks as follows:
\begin{enumerate}
	\item The platform commits to the commission rate $\alpha$ and the recommender policy.
	\item The sellers privately observes $q_A, q_B$ and decide whether to disclose it to the platform.
	\item The sellers simultaneously set prices.
	\item The platform observes prices and whether the sellers disclosed quality, and recommends sellers to customers. After that, the payoffs are realized and the game ends.
\end{enumerate}

\section{Results}
Firstly, I consider the case with observable qualities and without the recommender system. I need this comparison in order to demonstrate how and via which mechanisms the presence of the recommender system affects the equilibrium behavior of sellers.

@@

	 Given the demand, 	sellers set prices $p_A, p_B$ simultaneously to solve the following problems:
	 $$(1-\alpha)\cdot p_A \cdot \hat{x} = p_A \left(\frac{1}{2} - \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} + \frac{q_A - q_B}{2}\right) \to \underset{p_A}{\max}$$
	 $$(1-\alpha)\cdot p_B \cdot (1-\hat{x}) = p_B \left(\frac{1}{2} + \frac{\gamma - \alpha}{\gamma} \frac{p_A - p_B}{2} - \frac{q_A - q_B}{2}\right) \to \underset{p_B}{\max}$$
	 The system of first order conditions is given by:
	 \begin{align*}
	 \begin{cases}
	 p_A = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_B}{2} + \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}\\
	 p_B = \frac{\gamma}{2(\gamma - \alpha)} + \frac{p_A}{2} - \frac{\gamma(q_A - q_B)}{2(\gamma - \alpha)}
	 \end{cases}
	 \end{align*}
	 Solving the system above yields $p_i = \frac{\gamma}{\gamma - \alpha} + \frac{\gamma(q_i - q_{-i})}{3(\gamma - \alpha)}$
	\end{proof}
Comparing the results of the Proposition \ref{with} with Lemma \ref{without} several things are worth noting. First of all, with the recommender system the term $\frac{p_A-p_B}{2}$ in the expression for demand has an extra multiplier $\frac{\gamma-\alpha}{\gamma} \in (0, 1)$. The latter fact means that is the seller $A$ decides to undercut the competitor and sets a price $p_B - \varepsilon$, with the recommender system the resulting gain in demand will be $\frac{\gamma - \alpha}{\gamma} \frac{\varepsilon}{2}$ which is strictly below $\frac{\varepsilon}{2}$ the result we obtained for the case without the platform. That means that now the sellers have fewer incentives to undercut the competitor, i.e. the presence of a recommender system which has stakes in sellers' profits softens competition. As a result, the equilibrium prices a higher. The picture below (Fig. \ref{fig1}) demonstrates graphically the impact of the recommender system.

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.6\textwidth]{ImpactRS}
		\caption{Impact of RS $(p_A < p_B)$}\label{fig1}
	\end{figure}
	As we can see, there exists a share of customers $\Delta = \frac{\alpha}{2 \gamma} (p_B - p_A)$ such that they strictly prefer the firm $A$ to the firm $B$ while are recommended the firm $B$. This happens since from the platform's standpoint the losses in consumer surplus are lower than gains driven by high-price sales. Note, also, that $\frac{\partial \Delta}{\partial \alpha} > 0$, $\frac{\partial \Delta}{\partial \gamma} < 0$ and $\lim_{\gamma \to \infty} \Delta = 0$. That is, the impact of the platform eliminates completely if $\gamma$ becomes very high, and the more the platform is profit-oriented (the higher is $\alpha$) the higher is $\Delta$.
	\subsection{Qualities are not observable}
Here I consider the case when qualities $q_A, q_B$ are not observable to the platform unless the sellers decide to disclose it. In this subsection, I derive the optimal recommender system of the platform. Following \cite{levin2009quality} I consider symmetric PBE in which the sellers disclose the quality if and only if the quality falls above some threshold $q^*$ and do not disclose otherwise, and if the seller does not disclose, the beliefs of the platform regarding the quality do not depend on prices.

The optimal platform's policy can be found separately: firstly the platform may obtain the optimal recommendation (as a function of $\alpha$) treating the commission rate $\alpha$ as given, and then find an optimal commission rate. The next proposition characterizes the optimal recommendation strategy given the commission rate $\alpha$.

@@

From the Bayes rule, it follows that if the seller does not disclose the information the platform forms beliefs that the quality is $\frac{q^*}{2}$. Given the platform's beliefs regarding the quality, the proof exactly repeats that of Proposition \ref{with}. It remains to find $q^*$. This threshold is found as the minimal quality that renders the seller indifferent between disclosing and not disclosing the information. That is, $q^*$ is determined as a solution to the following equation	\begin{align*}

@@

However, it turns out that depending on the cost of information disclosure, the amount of information disclosed in equilibrium depends differently on the recommender system precision.
